\LARGE\singlespacing{\textbf{Supplementary Information} \\ \medskip Estimating \emph{Mycobacterium} tuberculosis transmission in a South African clinic: Spatiotemporal model based on person movements}

### 2024-05-21 doc/paper/supplements.tex

\title{\LARGE\singlespacing{\textbf{Supplementary Information} \\ \medskip
Estimating \emph{Mycobacterium} tuberculosis transmission in a South African clinic: Spatiotemporal model}}

@@

We reconnected tracks that most likely belonged to the same person. First, we automatically matched tracks that were less than 10\,s and 3\,m apart while moving or less than 60\,s and 0.5\,m apart while sitting. Second, we manually reconnected the remaining tracks if they were likely from the same person. For that purpose, we developed a Shiny app and a matching manual (\url{https://github.com/nbanho/clinic-transmission/wiki/Matching-Manual}) to assist the students performing the matching.

The Shiny app is shown in \Cref{fig:shiny-app}. In the centre panel, a floor plan of the clinic is shown in (x,y)-space. Different areas of the clinic are delineated by colours, \eg seating areas are marked in purple and exit areas are marked in blue. Tracks are shown as lines with a open and crossed circles indicating the tracks' start and end, respectively. The currently selected track is shown in black, possible links in colours, and alternatives for the selected possible link in light grey.

@@

The Shiny app's left panel shows descriptive information about the loaded file, the date of the tracking data, the initial number of IDs before any manual linkages, the current number of IDs after considering manual linkages, and the current number of made linkages. It further shows the current Pid with shortcuts to move to the next or previous Pid, or the next or previous Pid at the entrance. Linkages can be done either forward or backward in time. The total duration of the Pid track (time in the clinic) is shown together with the number of connected tracks within Pid. The sliding bars allow filtering of possible links (Oid) according to difference in time, distance, or height, with quick filters to switch between common thresholds. The first dropdown menu selects the Oid for which alternative matches should be shown in the floor plan. The idea is to prevent false linkages when there are more plausible alternatives for the Oid. A connection is made by selecting the Oid in the "Link with IDs" dropdown menu and then clicking "Link IDs". After that, the app updates the floor plan, matching table, and descriptive information. Erroneous link can be revoked by clicking "Unlink last ID", if the last link in time should be revoked, or by clicking "Unlink first ID", if the first link in time should be revoked. If the Pid could not be connected further, its track can be finished by entering a label and clicking "Label/End track". The person-tracking data is then updated with the reconnected track and saved automatically to the file shown at the bottom.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics[width=\linewidth]{doc/paper/shiny-tool-screenshot.png}
    \caption{Screenshot of the Shiny app used for manual linking tracking IDs that probably belong together.}
    \label{fig:shiny-app}
\end{figure}

We obtained 3,319 tracks after preprocessing. Further exclusion of short tracks <5\,min resulted in 1,563 tracks used for analysis: 209 (13\%) labeled as "Clean", 125 (8\%) as "Staff", 213 (14\%) as "Lost enter", 246 (16\%) as "Lost exit", and 770 (49\%) as "Lost both". A sample of person tracks are shown in \Cref{fig:tracking-examples}.

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/data/example-patient-tracks.png}
    \caption{Sample of person tracks from our video sensors after preprocessing.}
    \label{fig:tracking-examples}
\end{figure}

@@

The diffusion constant $D$ is unknown. We approximate it using an empirical relationship between the eddy diffusion coefficient $K$ ($K \approx D$ in m$^2$\,s$^{-1}$) and the outdoor air exchange rate $AER$ for CO$_2$ particles\cite{Cheng2011EnvSciTech,Foat2020BE}
\begin{align}
    K = (0.52 \cdot AER + 8.61\cdot10^{-5} \cdot \text{s}^{-1}) \cdot V^{\frac{2}{3}},
\end{align}
where $V$ is the volume of the airspace (in m$^3$). We thus assume that pathogen-carrying particles spread with a similar speed than CO$_2$ particles, which likely overestimates the diffusion constant, because CO$_2$ particles are smaller than \emph{Mtb}-carrying particles.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/inputs/quanta-generation-rate.png}
    \caption{Prior distribution (log scale) for the quanta generation rate (quanta h$^{-1}$) by activity level (sitting and walking), along with reported average estimates from the literature\cite{Andrews2014JID,Riley1962ARRD,Escombe2008PLoSMed,Nardell1991ARRD,Dharmadhikari2012AJRCCM}.}
    \label{fig:quanta-distribution}
\end{figure}

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/inputs/bacterial-inactivation-rate.png}
    \caption{Prior distribution (log scale) for the bacterial inactivation rate (quanta h$^{-1}$) of \emph{Mtb} along with reported estimates from the literature\cite{Loudon1969AMRRD,Gannon2007ResVetSci,Klein2014IJMyco,Lever2000LettersAppliedMicrobio}.}
    \label{fig:lambda-distribution}
\end{figure}

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/inputs/undiagnosed-tb-patients.png}
    \caption{Counts (\%) of the daily number of diagnosed TB patients visiting the clinic according to clinical records. The number of diagnosed patients is modeled with a Multinomial distribution based on these counts.}
    \label{fig:undiagnosed-distribution}
\end{figure}

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/modeling/mean-roi-assumptions.png}
    \caption{Mean risk of infection per clinic attendee if \textbf{(a)}~the number of infectious people in the clinic corresponded to the prevalence of TB in the population, or \textbf{(b)}~the number of infectious people in the clinic corresponded to the number of bacteriologically confirmed TB patients and those suspected of having TB due to respiratory symptoms when attending the clinic. Low: <0.1\%, Medium: 0.1-1\%, High: >1\% risk of infection.}
    \label{fig:assumptions-results}
\end{figure}

>>>>

\title{\LARGE\singlespacing{\textbf{Supplementary Information} \\ \medskip
Estimating \emph{Mycobacterium} tuberculosis transmission in a South African clinic: Spatiotemporal model based on person movements}}

@@

We reconnected tracks that most likely belonged to the same person. First, we automatically matched tracks that were less than 10\,s and 3\,m apart while moving or less than 60\,s and 0.5\,m apart while sitting. Second, we manually reconnected the remaining tracks if they were likely from the same person. For that purpose, we developed a Shiny application and a matching manual (\url{https://github.com/nbanho/clinic-transmission/wiki/Matching-Manual}) to assist the students performing the matching.

The Shiny application is shown in \Cref{fig:shiny-app}. In the centre panel, a floor plan of the clinic is shown in (x,y)-space. Different areas of the clinic are delineated by colours, \eg seating areas are marked in purple and exit areas are marked in blue. Tracks are shown as lines with a open and crossed circles indicating the tracks' start and end, respectively. The currently selected track is shown in black, possible links in colours, and alternatives for the selected possible link in light grey.

@@

The Shiny application's left panel shows descriptive information about the loaded file, the date of the tracking data, the initial number of IDs before any manual linkages, the current number of IDs after considering manual linkages, and the current number of made linkages. It further shows the current Pid with shortcuts to move to the next or previous Pid, or the next or previous Pid at the entrance. Linkages can be done either forward or backward in time. The total duration of the Pid track (time in the clinic) is shown together with the number of connected tracks within Pid. The sliding bars allow filtering of possible links (Oid) according to difference in time, distance, or height, with quick filters to switch between common thresholds. The first dropdown menu selects the Oid for which alternative matches should be shown in the floor plan. The idea is to prevent false linkages when there are more plausible alternatives for the Oid. A connection is made by selecting the Oid in the "Link with IDs" dropdown menu and then clicking "Link IDs". After that, the app updates the floor plan, matching table, and descriptive information. Erroneous link can be revoked by clicking "Unlink last ID", if the last link in time should be revoked, or by clicking "Unlink first ID", if the first link in time should be revoked. If the Pid could not be connected further, its track can be finished by entering a label and clicking "Label/End track". The person-tracking data is then updated with the reconnected track and saved automatically to the file shown at the bottom.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics[width=\linewidth]{doc/paper/shiny-tool-screenshot.png}
    \caption{Screenshot of the Shiny application used for reconnecting interrupted tracking IDs.}
    \label{fig:shiny-app}
\end{figure}

We obtained 3,319 tracks after preprocessing. Further exclusion of short tracks <5\,min resulted in 1,563 tracks used for analysis: 209 (13\%) labeled as "Clean", 125 (8\%) as "Staff", 213 (14\%) as "Lost enter", 246 (16\%) as "Lost exit", and 770 (49\%) as "Lost both".

@@

The diffusion constant $D$ is unknown. We approximate it using an empirical relationship between the eddy diffusion coefficient $K$ ($K \approx D$ in m$^2$\,s$^{-1}$) and the outdoor air exchange rate $AER$ for CO$_2$ particles\cite{Cheng2011EnvSciTech,Foat2020BE}
\begin{align}
    K = (0.52 \cdot AER + 8.61\cdot10^{-5} \cdot \text{s}^{-1}) \cdot V^{\frac{2}{3}},
\end{align}
where $V$ is the volume of the airspace (in m$^3$). We thus assume that pathogen-carrying particles spread with a similar speed than CO$_2$ particles, which may overestimate the diffusion constant because CO$_2$ particles are smaller than \emph{Mtb}-carrying particles.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/inputs/quanta-generation-rate.png}
    \caption[Prior distribution for the quanta generation rate]{Prior distribution (log scale) for the quanta generation rate (quanta h$^{-1}$) by activity level (sitting and walking), along with reported average estimates from the literature\cite{Andrews2014JID,Riley1962ARRD,Escombe2008PLoSMed,Nardell1991ARRD,Dharmadhikari2012AJRCCM}.}
    \label{fig:quanta-distribution}
\end{figure}

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/inputs/bacterial-inactivation-rate.png}
    \caption[Prior distribution for the bacterial inactivation rate]{Prior distribution (log scale) for the bacterial inactivation rate (quanta h$^{-1}$) of \emph{Mtb} along with reported estimates from the literature\cite{Loudon1969AMRRD,Gannon2007ResVetSci,Klein2014IJMyco,Lever2000LettersAppliedMicrobio}.}
    \label{fig:lambda-distribution}
\end{figure}

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/inputs/undiagnosed-tb-patients.png}
    \caption[Counts of the daily number of diagnosed TB patients visiting the clinic]{Counts (\%) of the daily number of diagnosed TB patients visiting the clinic. The number of diagnosed patients is modeled with a Multinomial distribution based on these counts.}
    \label{fig:undiagnosed-distribution}
\end{figure}

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/modeling/mean-roi-assumptions.png}
    \caption[Mean risk of infection per clinic attendee when varying assumptions about the number of infectious people in the clinic]{Mean risk of infection per clinic attendee if \textbf{(a)}~the number of infectious people in the clinic corresponded to the prevalence of TB in the population, or \textbf{(b)}~the number of infectious people in the clinic corresponded to the number of bacteriologically confirmed TB patients and those suspected of having TB due to respiratory symptoms when attending the clinic. Low: <0.1\%, Medium: 0.1-1\%, High: >1\% risk of infection.}
    \label{fig:assumptions-results}
\end{figure}
